### 2018-02-11 writeup/Outline/tex/sections/07-discussion.tex

\section{Discussion}

\begin{enumerate}
  \item Using the true generating function under-states our uncertainty -- can connect to the idea that using model chaining also under-states our uncertainty
  \item These are idealized scenarios but are motivated by real-world problems in flood adaptation (begin to build theory which is applicable to old cities in U.S., fast-growing cities around the world, and more)
  \item More carefully critique the current literature in light of these results
  \item Future steps
  \begin{enumerate}
    \item Better loss estimation
    \item Real options model
    \item Better ways to downscale climate projections for estimating flood loss
    \item Extend beyond annual-maximum
  \end{enumerate}
\end{enumerate}