### 2018-10-30 poster-nexus-conference/blocks/climate-change.tex

\begin{block}{Secular Change}
  Risk, defined as the product of \emph{hazard} and \emph{exposure}, is changing (``nonstationary''):
  \begin{itemize}
    \item \gls{acc} affects the intensity, location, and frequency of hydroclimate extremes \cite{Held:2006jo, OGorman:2009hj}
    \item changes in land use, river channels, and water use affect local hydrologic cycles \cite{Merz:2014gf}
    \item urbanization and development drive increasing exposure to floods \cite{Jongman:2012cr} and hurricanes \cite{Peduzzi:2012iq}
  \end{itemize}
\end{block}