### 2018-10-30 poster-nexus-conference/blocks/observed-lfv.tex

\begin{block}{Observed \acrshort{lfv}}
  Our analysis is motivated by analysis of historical and paleo records of hydroclimate systems, which often show key modes of variability on interannual to multidecadal time scales \cite{Merz:2014gf, Swierczynski:2012km}.
  The following time series and wavelet global spectra are representative of this \gls{lfv}.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{observed-lfv.pdf}
    \caption{
      Hydroclimate time series vary on many time scales.
      (a) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
      (b) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda} \cite{Cook:2010bz}.
      (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
      (d) like (c), for the \gls{lbda} data.
    }\label{fig:observed-lfv}
  \end{figure}
\end{block}