### 2018-10-30 poster-nexus-conference/blocks/uncertainty.tex

\begin{block}{Types of Uncertainty}
    \begin{figure}
        \centering
        \caption{
            A stylized illustration of (a) irreducible and (b) estimation uncertainty.
        }\label{fig:conceptual-sketch}
        \includegraphics[width=\textwidth]{conceptual-sketch.pdf}
    \end{figure}
    \begin{description}
        \item[Irreducible uncertainty] cannot be resolved with better models or data and is dominated in the short term by chaotic behavior of the climate, and in the long term by the uncertainty in future \acrlong{acc} (\cref{fig:conceptual-sketch}a) 
        \item[estimation uncertainty:] the length of a historical record limits the potential to identify different climate signals (\cref{fig:conceptual-sketch}b)
    \end{description}
\end{block}