### 2018-11-12 poster-nexus-conference/blocks/conclusions.tex

{   
    \large
    \setbeamercolor*{block body}{fg=black,bg=wcblue!30}
    \begin{block}{Key Findings}
        Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future:
        \begin{itemize}
            \item If uncertainty increases into future then adaptation strategies with short planning periods are preferred
            \item Risk profile of short-$M$ projects dominated by \acrlong{lfv} \cite{Jain:2001hz, Hodgkins:2017hw}
            \item Even though the \gls{hmm} is an imperfect analog for \gls{enso}, it performs well for short planning periods
            \item When the planning period is long, trends must be estimated explicitly (requiring more data)
        \end{itemize}
    \end{block}
}