### 2018-11-12 poster-nexus-conference/blocks/systematic-experiments.tex

\begin{block}{Experiment Results}
  We evaluate bias and variance as a function of $M$, $N$, and estimation method for three scenarios.
  \begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{secular-only-nino3-bias-variance.pdf}
    \caption{
      Expected estimation bias and variance for sequences with secular change only (no \gls{lfv}).
    }\label{fig:secular-nino3-bias-variance}
  \end{figure}
  \begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{lfv-only-nino3-bias-variance.pdf}
    \caption{
      As \cref{fig:secular-nino3-bias-variance} but for sequences generated with zero secular change and \gls{lfv} from the \gls{enso} index.
    }\label{fig:lfv-nino3-bias-variance}
  \end{figure}
  \begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{lfv-secular-nino3-bias-variance.pdf}
    \caption{
      As \cref{fig:secular-nino3-bias-variance} but for sequences generated from the NINO3 model with both \gls{lfv} and secular change.
    }\label{fig:lfv-secular-nino3-bias-variance}
  \end{figure}
\end{block}