### 2019-05-17 report-earths-future/reviewer-response.md
Final version

# Response to Reviewers

We again thank Dr. Mari, Dr. Montanari, and one anonymous reviewer for their helpful comments on our paper "Robust Adaptation to Multi-Scale Climate Variability".
Please see below our responses to the minor comments from reviewer #2.

I am glad to see you added a discussion about robustness. To reflect the focus of the article, which is on sources of uncertainty, I would reword title as something like, "Understanding Multi-Scale Climate Variability for Robust Adaption."

> Thank you for your helpful comments. The two titles are very similar, but we feel that the current wording is the most understandable and direct. In particular, we think that the title as-is focuses not about understanding multiscale climate variability in the context of robust adaptation, but how to think abut adaptation given the multiscale variability that exists.

Line 62-63: I appreciate that you reworded this sentence. Minor suggestion: "In this paper we focus more narrowly on the temporal ..." (to avoid using the word question twice in this sentence).

> We thank the reviewer for this suggestion and agree that this wording is clearer. The updated wording is on L63.

The introduction of the Hurst phenomenon is an interesting addition to your paper. I suggest being more upfront with a definition, e.g.: The Hurst phenomenon is a well-known mathematical relationship between the magnitude of natural events and their standard deviation of occurrence, ...

> Thank you for this suggestion. We have re-worded this sentence to include a bit more information about the Hurst phenomenon. It how reads "The Hurst phenomenon is a well known mathematical relationship which describes the long memory of processes found in in geophysics, physics, biology, medicine, traffic, network dynamics, and finance," (L121).

Lines 285-291: I think these sentences would be better presented as an equation.

> Thank you for this suggestion. We agree that describing these operations in a paragraph is rather convoluted. However, describing these operations using equations would actually require several equations: to define the estimated threshold, the bias, and the variance. We feel that adding these equations would interrupt the flow of this paragraph and add needless complexity to a paper which already asks the reader to keep track of a lot of ideas at once. The openly accessible codes make it easy for anyone who wishes to reproduce or modify this work, and so we would like to leave these sentences as they are.

I'm confused by figure 3. The variables N and M could be redefined in the caption. Isn't M the investment lifespan, i.e. occurring in the future, and N is the historical sample period; if so, why does the third column show estimated and observed exceedances for N and M, respectively?

> Thanks for pointing this out. We have updated figure 3 and its caption to improve clarity. Specifically, we have re-worded the box which read "Estimated Exceedances per Year" to "Projected Future Exceedances per Year" to emphasize that the estimating model and N-year historical record are used to estimate future risk.

Line 417: Reword sentence to reflect the more limited illustration in fig. 2.

> We agree that based on figure 2 we should restrict our focus to climate uncertainties, rather than initiating a discussion of socio-economic and other uncertainties. We have re-worded this sentence to read  "As illustrated schematically in fig 2, large uncertainties (e.g., as to future CO2 concentrations and local climate impacts) lead to large intrinsic uncertainty in projections of future climate risk." (L419).